---
title: "Assignment 3"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---


#### 1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the variable z and print the value stored in each z.

  a. x^a^^^b^^
  b. (x^a^)^b^
  c. 3x^3^+2x^2^+1

```{r}
# a. 
x = 1.1; a = 2.2; b = 3.3
z <- x^a^b
print(z)

# b.  
z <- (x^a)^b
print(z)

# c. 
z <- 3*x^3 + 2*x^2 + 1
print(z)
```


#### 2. Using the rep and seq functions, create the following vectors:

  a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
  b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
  c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
# a. 
z <- c(seq(from=1, to=7), seq(from=8, to=1))
print(z)

# b. 
z <- rep(1:5, c(1,2,3,4,5))
print(z)

# c.
z <- rep(seq(from=5, to=1), c(1,2,3,4,5))
print(z)
```


#### 3. Create a vector using the code: queue <- c("sheep", "fox", "owl", "ant"), where queue represents the animals that are lined up to enter an exclusive club, with the sheep at the front of the line. Using R expressions, update queue as:

  a. the serpent arrives and gets in the back of the line;
  b. the sheep enters the club (so disappears from the line);
  c. the donkey arrives and talks his way to the front of the line;
  d. the serpent gets impatient and leaves;
  e. the aphid arrives and the ant invites him to cut in line (hint: check out the append function).
  f. Finally, determine the position of the aphid in the line (hint: use the which function).
  
```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)

# a. 
queue <- c(queue[1:4], "serpent")
print(queue)

# b. 
queue <- queue[-c(1)]
print(queue)

# c. 
queue <- c("donkey", queue[1:4])
print(queue)

# d. 
queue <- queue[-c(5)]
print(queue)

# e. 
queue <- append(queue, "aphid", after = 3)
print(queue)

# f. 
which("aphid" == queue)[[1]]
```
