---
title: "Programming in R"
author: "Emily K. Power"
date: "12 January 2023"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
---

### Advantages of R  
  * interactive use  
  * graphics, statistics  
  * very active community of contributors  
  * works on multiple platforms  

### Disadvantages of R  
  * lazy evaluation  
  * some packages are poorly documented  
  * functions are hard to learn  
  * unreliable packages  
  * problems with big data (multiple GBs)  

### Let's start with the basics  

### Assignment operator: used to assign a new value to a variable (<-)
```{r}
x <- 5      # displays values in environment
print(x)    # prints output of x to console
x           # same as print(x), can also just highlight x and run

y = 4           # legal, but not used except in function arguments
print (y)
y = y + 1.1     # 5.1; takes previous value of y (4) and adds 1.1
y <- y + 1.1    # 6.2; makes more sense than =
```

### To run code all at once:
  * highlight everything and run
  * click 'source' button

### Variables: used to store information like a container
  * can hold many types of data
```{r}
z <- 3
plantHeight <- 10  # 'camel case format' (1 lowercase, 2 uppercase)
plant_height <- 3.3  # 'snake case format', a preferred method
plant.height <- 4.2  # generally avoid
. <- 5.5 # reserve for temporary variable (. acts as placeholder)
```

### Functions: blocks of code that perform a task
  * can use a short command over again instead of writing block of code multiple times
  * you can create your own function!
```{r}
square <- function(x = NULL){
  x <- x^2  # take variable x, square it and override initial x
  print(x)
}
## put argument in ()

square(x=3)

z <- 103
square(x = z)  # argument name is x
```

### Built In Functions
```{r}
sum(109, 3, 10)  
## look at package info using ?sum or going to Help panel
```

### Atomic Vectors
  * one dimensional (a single row)
  * one of the most fundamental data structures in R programming

### Types of Atomic Vectors:
  * character strings (usually within quotes)
  * integers (whole numbers; numeric)
  * double (real numbers, decimals; numeric)
  * logical (TRUE or FALSE)
  * factor (categorizes and groups variable)

### c Function - combines different elements & puts as one vector
  * useful for creating vectors
```{r}
z <- c(3.2, 5, 5, 6)
print(z)
class(z)          # function that returns class of variable
typeof(z)         # function that returns type of variable
is.numeric(z)     # function that returns T/F
## this is TRUE because z is numeric (asks 'is z numeric'?)
## if you wrote is.logical(z), it would return FALSE

# c() always 'flattens' to a vector
z <- c(c(3,4), c(5,6))   # still works!
```

### Character Vectors
```{r}
z <- c("perch", "bass", "trout")
print(z)
z <- c("This is only 'one' character string", "a second", "a third")
print(z)
typeof(z)
is.character(z)  # is. functions fest whether it is a certain data type
```

### Logical (Boolean), no quotes, all caps: TRUE and/or FALSE
```{r}
z <- c(TRUE, FALSE, T, F)  # as. functions coerce data type
z <- as.character(z)
is.logical(z)
```

### Type (vector property)
  * is.numeric / as.character
  * typeof()

### Length (vector property)
```{r}
length(z)  # number of items in function
dim(z)     # will return 'NULL' because vectors only have 1 dimension
```

### Names
```{r}
z <- runif(5)  # function gives 5 random uniform numbers between 0 & 1
names(z)       # no names
names(z) <- c("chow", "pug", "beagle", "greyhound", "akita") # add names to random uniform values
print(z)
names(z)  # shows names now
names(z) <- NULL  # shows no names
```

### NA values - missing data
```{r}
z <- c(3.2, 3.5, NA)  # NA could be an empty cell in excel
typeof(z)
mean(z)          # if there are any NAs in vector, arithmetic won't work
anyNA(z)         # checks if there are any NAs in vector
is.na(z)         # shows which cells in vector are NAs
which(is.na(z))  # tells you where the NA is (which specific index)
## helpful for exploring data and determining where NAs are
```

### Subsetting vectors - filters for certain elements
  * gives set of conditions of what we want from vector
```{r}
z <- c(3.1, 9.2, 1.3, 0.4, 7.5)
z[4]         # square brackets subset by index (run the 4th element)
z[c(4,5)]    # use c to return multiple indices
z[-c(2,3)]   # '-' sign excludes elements listed in parentheses
z[-1]        # returns everything except first value
z[z==7.5]    # '==' only returns values that meet the condition (= 7.5)
## '==' means 'exact match'
z[z<7.5]     # 'return all elements in 7 that are less than 7.5'
## use logical statements within square brackets to subset by conditions

which(z < 7.5)  # 'which elements in z are less than 7.5?'
```

### Creating logical vector
```{r}
z < 7.5  # returns location of elements less than 7.5
```

### Subsetting characters (named elements)
```{r}
names(z) <- c("a", "b", "c", "d", "e")  # names elements in z vector
z[c("a", "d")]  # returns a & d and their corresponding values
```

### Subset function - alternative to using brackets
```{r}
subset(x = z, subset = z > 1.5)  
## first argument is what you want to subset
  ## x is automatically built into function
## subset = condition you want
```

### Randomly Sample
  * sample function
```{r}
story_vec <- c("A", "Frog", "Jumped", "Here")
sample(story_vec)
```

### randomly select 3 elements from vector
```{r}
sample(story_vec, size = 3)

story_vec <- c(story_vec[1], "Green", story_vec[2:4])
## adds 'green' to the story between 1st and 2nd elements
## : signifies a range between 1st and last numbers

story_vec[2] <- "Blue"
## replaces 2nd element (green)
```

### vector function
```{r}
vec <- vector(mode = "numeric", length = 5)
```

### rep and seq functions - make patterns of values
```{r}
z <- rep(x = 0, times = 100)  # repeats value 0 100 times
z <- rep(x = 1:4, each = 3)
## repeats each element 3 times

z <- seq(from = 2, to = 4)
## returns sequence of numbers from 2 to 4
seq(from = 2, to = 4, by = 0.5)
## returns sequence of numbers from 2.4 by units of 0.5
z <- runif(5)
seq(from = 1, to = length(z))
## returns sequence of numbers for length of z if z is 5 elements
```