---
title: "Lecture 8: Loading in Data"
author: "Emily K. Power"
date: "26 January 2023"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
---

### Create and save a data set:
```{r}
# write.table(x=varName, file="outputFileName.csv", header=TRUE, sep=",")
```


### These functions read in a data set:
```{r}
# read.table(file="path/to/data.csv", header=TRUE, sep = ",")
# read.csv(file="data.csv", header=TRUE)
```


### Use RDS object when only working in R
  * helpful with large datasets
```{r}
# saveRDS(my_data, file="FileName.RDS")
# readRDS("FileName.RDS")
# p <- readRDS("FileName.RDS")
```

### Long vs. Wide data formats
  * **wide format**: contains values that do not repeat in the ID column
  * **long format**: contains values that DO repeat in the ID column
```{r}
library(tidyverse)
glimpse(billboard)  # wide format, many columns
b1 <- billboard %>%
  pivot_longer(
    cols = starts_with("wk"), # cols specifies which columns you want to make 'longer'
    names_to = "Week", # names_to creates name of new column which will contain the header names
    values_to = "Rank", # values_to specifies name of new column which will contain the values
    values_drop_na = TRUE  # removes any rows where the value = NA
  )

head(b1)
```

### pivot_wider
  * best for making occupancy matrix
```{r}
glimpse(fish_encounters)

fish_encounters %>%
  pivot_wider(names_from = station,  # which column you want to turn into multiple columns
              values_from = seen)  # which column contains the values for new column cells

fish_encounters %>%
  pivot_wider(names_from = station,
              values_from = seen, 
              values_fill = 0)
  # good format for occupancy matrix
```

### Dryad: makes research data freely reusable, citeable, and discoverable
```{r}
# type file.choose() into console to get path to data
dryadData <- read.table(file="/Users/emilypower/Desktop/BIOL_1007A/Data/NE_amphibians.csv", header=TRUE, sep=",")
glimpse(dryadData)
head(dryadData)
table(dryadData$species)  # allows you to see different groups of character column
summary(dryadData$mean.hydro)

str(dryadData$species)
dryadData$species <- factor (dryadData$species, labels=c("Spotted Salamander", "Wood Frog"))  # creating 'labels' to use for the plot
str(dryadData$species)
```

