---
title: "Entering the tidyverse (dplyr)"
author: "Emily K. Power"
date: "23 January 2023"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
---

### tidyverse: collection of opinionated packages that shares philosophy, grammar (how the code is structured), and data structures
  * eg. ggplot, dplyer

### Operators: symbols that tell R to perform different operations (between variables, functions, etc.)
  * **Arithmetic operators:** +, -, *, /, ^, ~
  * **Assignment operator:** <-
  * **Logical operators:** !(not), &(and), |(or)
  * **Relational operators:** ==, !=, >, <, >=, <=
  * **Miscellaneous operators:** %>%(forward pipe operator), %in%

### you only need to install packages once
  * install.packages("") in console
```{r}
library(tidyverse)  # library function loads in packages
```

### dplyr: new(er) package, provides a set of tools for manipulating data sets
  * specifically written to be fast
  * individual functions that correspond to common operations

### The core verbs:
  * **filter():** filters rows
  * **arrange():** arranges rows
  * **select():** selects columns
  * **group_by() and summarize():** manipulates data and summarizes stats
  * **mutate():** creates new variables with data frame

### built-in data set
```{r}
data(starwars)
class(starwars)
```

### Tibble (tbl): modern take on data frames
  * keeps great aspects of dfs, drops frustrating ones (change variables, change input type...)
```{r}
glimpse(starwars)  # much cleaner than str(starwars)
```

### NAs
```{r}
anyNA(starwars)  # are there any NAs in this dataset?
  # could also use is.na, complete.cases
starwarsClean <- starwars[complete.cases(starwars[,1:10]),]  # gets rid of all rows with NAs in the first 10 columns
anyNA(starwarsClean[,1:10])  # no NAs left
```

### filter(): picks/subsets observations (ROWS) by their values
```{r}
filter(starwarsClean, gender == "masculine" & height < 180)  # , also means &
filter(starwarsClean, gender == "masculine", height < 180, height > 100)  # can have multiple conditions for the same variable
```

### %in% operator (matching): similar to == but you can compare vectors of different lengths
```{r}
### sequence of letters
a <- LETTERS[1:10]
length(a)  # length of vector

b <- LETTERS[4:10]
length(b)

## output of %in% depends on whatever the first vector is
a %in% b  # gives logical statement, which elements of a are in b?
b %in% a  # which elements of b are in a?

## use %in% to subset
eyes <- filter(starwars, eye_color %in% c("blue", "brown"))
```

### arrange(): reorders rows
```{r}
arrange(starwarsClean, by=height)  # default is ascending order\
  # can use helper function desc() to make it descending
arrange(starwarsClean, by=desc(height))
arrange(starwarsClean, height, desc(mass))  # second variable used to break ties

sw <- arrange(starwars, by=height)
tail(sw)  # all NAs are at the end
```

### select(): chooses variables (COLUMNS) by their names
```{r}
select(starwarsClean, 1:10)
select(starwarsClean, name:species)
select(starwarsClean, -(films:starships))  # excludes those columns

## rearrange columns
select(starwarsClean, name, gender, species, everything())  # everything puts all other columns at the end
  # everything() helper function: useful if you have a couple variables you want to move to the beginning

## contains() helper function
select(starwarsClean, contains("color"))
  # other helper functions include:
    # ends_with()
    # starts_with()
    # num_range()  # numeric range

# select can also rename columns
select(starwarsClean, haircolor = hair_color)  # haircolor is new name
  # returns only renamed column
rename(starwarsClean, haircolor = hair_color)
  # returns entire data frame
```

### mutate(): creates new variables using functions of existing variables
```{r}
# let's create a new column that is divided by mass
mutate(starwarsClean, ratio = height/mass)

starwars_lbs <- mutate(starwarsClean, mass_lbs = mass*2.2)
view(starwars_lbs)
starwars_lbs <- select(starwars_lbs, 1:3, mass_lbs, everything())
glimpse(starwars_lbs)  # brought it to the front using select

# transmute
transmute(starwarsClean, mass_lbs=mass*2.2)  # only returns mutated columns
transmute(starwarsClean, mass, mass_lbs=mass*2.2, height)
```

### group_by() and summarize()
```{r}
summarize(starwarsClean, meanHeight = mean(height))  # throws NA if any NAs are in df - need to use na.rm

summarize(starwarsClean, meanHeight = mean(height), TotalNumber = n())

## use group_by for maximum usefulness
starwarsGenders <- group_by(starwars, gender)
head(starwarsGenders)
summarize(starwarsGenders, meanHeight=mean(height, na.rm=TRUE), TotalNumber=n())
```

### Piping %>%: used to emphasize a sequence of actions
  * allows you to pass an intermediate results onto the next function (uses output of one function as input of the next function)
  * avoid if you need to manipulate more than one object/variable at a time; or if variable is meaningful
  * formatting: should have a space before the %>% followed by new line
```{r}
starwarsClean %>%
  group_by(gender) %>%
  summarize(meanHeight=mean(height, na.rm=TRUE), TotalNumber=n())
  # much cleaner, don't have to use intermediate variables
```

### case_when() is useful for multiple if/ifelse statements
```{r}
starwarsClean %>%
  mutate(sp = case_when(species== "Human" ~ "Human", TRUE ~ "Non-Human"))
  # uses condition, puts "Human" if True in sp column, puts "Non-Human" if it's FALSE
```