---
title: "Finishing up Matrices and Data Frames"
author: "Emily K. Power"
date: "19 January 2023"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
---

```{r}
m <- matrix(data = 1:12, nrow = 3)
```

### subsetting based on elements
```{r}
m[1:2, ]
m[, 2:4]
```

### subset with logical (conditional) statements
```{r}
## select all columns for which the totals are greater than 15
colSums(m) > 15
m[, colSums(m) > 15]  # subsets based on conditions (sum of column > 15)
```

### row sums 
```{r}
## select rows that sum up to 22
m[rowSums(m)==22,]
m[rowSums(m)!=22,]
```

### LOGICAL OPERATORS:
  * == equals
  * != does not equal
  * > greater than 
  * < less than 

### subsetting to a vector changes the data type
```{r}
z <- m[1, ]
print(z)
str(z)

z2 <- m[1, , drop=FALSE]  # argument maintains the matrix identity
```

### simulate new matrix
```{r}
m2 <- matrix(data = runif(9), nrow=3)
m2[3,2]
```

### use assignment operator to substitute values
```{r}
m2[m2 > 0.6] <- NA  # makes any value >0.6 an NA

data <- iris
head(data)
tail(data)

data[3,2]  # numbered indices work, but sometimes it's better to use names
dataSub <- data[,c("Species", "Petal.Length")]
str(dataSub)
```

### sort a data frame by values
```{r}
orderedIris <- iris[order(iris$Petal.Length),]
```


## FUNCTIONS in R

### everything in R is a function
```{r}
sum(3,2)  # sum
3+2  # + sign is a function
sd
```

### User-defined functions
```{r}
# functionName <- function(argX=defaultX, argY=defaultY){
  ## curly bracket starts the body of the function
    ## lines of code
    ## notes
    ## create local variables - only 'visible' to R WITHIN the function
  ## always end with return(z) (like print())
# }

myFunc <- function(a=3, b=4){
  z <- a + b
  return(z)
}
myFunc()  # runs default arguments
myFunc(a=100, b=3.4)  # runs new values

## example of a bad function:
#myFuncBad <- function(a=3){
  #z <- a + b
  #return(z)
#}
# myFuncBad()  # error: object B not found

## how to use with no default values:
myFuncNull <- function(a=NULL, b=NULL){
  z <- a + b
  return(z)
}
#myFuncBad()  # returns 'integer 0'
```

### Multiple return statements
```{r}
########################################################################
# FUNCTION: HardyWeinberg
# input: all allele frequency p (0,1)
# output: p and the frequencies of 3 genotypes AA AB BB
#-----------------------------------------------------------------------
HardyWeinberg <- function(p = runif(1)){
  if(p > 1.0 | p < 0.0){
    return("Function failure: p must be between 0 and 1")
  }               ## | means 'or', {} means 'then'
  q <- 1 - p
  fAA <- p^2      ## 'frequency of AA allele is p^2
  fAB <- 2*p*q
  fBB <- q^2
  vecOut <- signif(c(p=p, AA=fAA, AB=fAB, BB=fBB), digits=3)
  return(vecOut)  ## signif outputs significant figures
}
########################################################################

HardyWeinberg()
freqs <- HardyWeinberg()
freqs
HardyWeinberg(p=3)  ## prints return message
```

### Create a complex default value
```{r}
########################################################################
# FUNCTION: fitLinear2
# fits simple regression line
# input: numeric vector (p) of predictor (x) and response (y)
# outputs: slope and p-value
#-----------------------------------------------------------------------
fitLinear2 <- function(p=NULL){
  if(is.null(p)){
    p <- list(x=runif(20), y=runif(20))
  }       # 'if p is null, then we'll make a variable p'
  myMod <- lm(p$x~p$y)
  myOut <- c(slope = summary(myMod)$coefficients[2,1], 
             pValue = summary(myMod)$coefficients[2,4])
  plot(x=p$x, y=p$y)  # quick plot to check output
  return(myOut)
}
########################################################################
fitLinear2()
myPars <- list(x=1:10, y=runif(10))
fitLinear2(p=myPars)
```