---
title: "Weekly Assignment 2"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---

## Creating Functions
1. Create a function that returns a data frame with 2 columns (named “Type” and “Value”) and 50 rows/observations.  
    * The first column should have “Control” as the first 25 observations and “Treatment” as the second half of observations.
    * The second column should have the first 25 values as random and normally distributed with a mean of 10 and standard deviation of 1.5. The next 25 values of the second column should be random and normally distributed with a mean of 45 and standard deviation of 2.
    * The arguments of your function should be the two means of the random normal numbers.
    * The body of the function can be a single line of code (if you want a challenge) or broken up into multiple code blocks. 
```{r}
########################################################################
# FUNCTION: new_function
# returns a data frame with 2 columns (named “Type” and “Value”) and 50 rows/observations
# input: mean and standard deviation
# outputs: "Type" and "Value"
#-----------------------------------------------------------------------
new_function <- function(means=c(10, 45), sds=c(1.5, 2), n=25){
  Type <- rep(c("Control", "Treatment"), each=n)  # create "Type" column with n "Control" and n "Treatment" variables
  Value <- c(rnorm(n, mean=means[1], sd=sds[1]), rnorm(n, mean=means[2], sd=sds[2]))  # generate random normally distributed values with set means and standard deviations
  df <- data.frame(Type, Value)  # generate data frame
  return(df)  # returns data frame
}
```



2. Save your new function’s output as a variable, and view the first 6 rows of the data frame.
```{r}
set.seed(123)  # guarantees that same random values are produced each time code is run
output <- new_function()  # saves function as variable
head(output)  # prints first 6 rows of data
```


3. Let’s say you would like to know whether there is a statistically significant difference in “Value” (response variable y) depending on “Type” (explanatory variable x). Type `?aov` in the console to determine how to run an analysis of variance (ANOVA) on your simulated data. The formula is written as `yVar~xVar` with yVar as the column name of your response variable and xVar as the column name of your explanatory variable. Write a line of code that displays a summary of your ANOVA.
```{r}
test <- aov(Value~Type, data = output)  # aov() function fits analysis of variance model by a call to 1m for each stratum
summary(test)  # prints test summary
```


4. Finally create a function that uses a data frame as its input and returns only the p-value of an ANOVA summary table (feel free to use Google/Stack Overflow). Write your code in such a way that you can use any simulated data set with two columns (i.e. the output of your first function) as your ANOVA function’s argument.
```{r}
########################################################################
# FUNCTION: ANOVA_function
# returns the p-value of an ANOVA summary table
# input: a data frame
# outputs: p-value
#-----------------------------------------------------------------------
ANOVA_function <- function(df=NULL){
  test <- aov(Value ~ Type, data=df)  # same as before
  test_summary <- summary(test)  # makes test summary
  p_value <- test_summary[[1]][["Pr(>F)"]][1]  # pulls p value from test summary data frame
  return(p_value)  # returns p value
}
```
```{r}
ANOVA_function(df=output)

## you can put your simulate data function in you p-value function
ANOVA_function(new_function())
```

