---
title: "Assignment 6"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---

## Data manipulations using the `dplyr` package
This homework assignment focus on data manipulation in `R`. Complete these problems using the `dplyr` and `tidyverse` packages.

1. Examine the structure of the iris data set. How many observations and variables are in the data set?
```{r}
library(dplyr)  # loads dplyr library
data(iris)  # loads iris data set
str(iris)  # or glimpse(iris), examines structure of data set
# 150 observations
# 5 variables
```


2. Create a new data frame `iris1` that contains only the species *virginica* and *versicolor* with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the data set?
```{r}
iris1 <- filter(iris, Species == "virginica" | Species == "versicolor", Sepal.Length > 6, Sepal.Width > 2.5)  # filter() function filters data set to only contain 2 species with specified sepal measurements
  # have to separate species (| - 'or' logical operator) because the vectors are different lengths
  # must use , instead of & because there are specified conditions prior
  # or iris1 <- filter(iris, Species %in% c("virginica", "versicolor"), Sepal.Length > 6, Sepal.Width > 2.5)
iris1  # prints variable
```


3. Now, create a `iris2` data frame from `iris1` that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the data set?
```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)  # select() function creates data frame with desired columns
iris2  # prints variable
# 28 observations
# 3 variables
```


4. Create an `iris3` data frame from `iris2` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this data set.
```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length))  # arrange() function orders iris2 data by specified variable (Sepal.Length with values in descending order)
head(iris3)  # prints first 6 rows of data frame
```


5. Create an `iris4` data frame from `iris3` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the data set?
```{r}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length * Sepal.Width)  # mutate() function adds a new column: Sepal.Area, which utilizes values from Sepal.Length and Sepal.Width
iris4  # prints variable
# 28 observations
# 4 variables
```


6. Create the variable `irisTab` that shows the average sepal length, the average sepal width, and the sample size of the entire `iris4` data frame and print `irisTab`.
```{r}
irisTab <- summarize(iris4, Mean.Sepal.Length = mean(Sepal.Length), Mean.Sepal.Width = mean(Sepal.Width), Sample.Size = n())  # summarize() function creates a summary table of mean values and sample size using data from iris4 
irisTab  # prints variable
```


7. Finally, create `iris5` that calculates the average sepal length, the average sepal width, and the sample size *for each species* of in the `iris4` data frame and print `iris5`.
```{r}
iris5 <- iris4 %>% # begins with values from iris4 dataframe 
  group_by(Species) %>% # groups iris4 data by species
  summarize(Mean.Sepal.Length = mean(Sepal.Length), Mean.Sepal.Width = mean(Sepal.Width), Sample.Size = n())  # summarizes values grouped by species
  # piping %>% limits usage of intermediate variables
iris5  # prints variable
```


8. In these exercises, you have successively modified different versions of the data frame `iris1 iris2 iris3 iris4 iris5`. At each stage, the output data frame from one operation serves as the input for the next. A more efficient way to do this is to use the pipe operator `%>%` from the `tidyr` package. Rework all of your previous statements (except for `irisTab`) into an extended piping operation that uses `iris` as the input and generates `irisFinal` as the output.
```{r}
irisFinal <- 
  iris %>%  # selects iris data set
  filter(Species == "virginica" | Species == "versicolor", Sepal.Length > 6, Sepal.Width > 2.5) %>%  # filters by species and specified sepal measurements
  select(Species, Sepal.Length, Sepal.Width) %>%  # selects 3 columns from filtered data set
  arrange(by=desc(Sepal.Length)) %>%  # arranges data by descending Sepal.Length values
  mutate(Sepal.Area = Sepal.Length * Sepal.Width) %>%  # creates Sepal.Area value
  group_by(Species) %>%  # groups data by species
  summarize(Mean.Sepal.Length = mean(Sepal.Length), Mean.Sepal.Width = mean(Sepal.Width), Sample.Size = n())  # creates summary table with mean values and sample size
irisFinal  # prints variable
```

