---
title: "Assignment 5"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---

## Continued from Assignment 4  

#### 2. Create a list with the following named elements:  

  * `my_matrix`, which is a 4 x 4 matrix filled with random uniform values
  * `my_logical`, which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it.
  * `my_letters`, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_matrix <- matrix(data = runif(16), nrow=4)
my_matrix

my_logical <- rnorm(100)
my_logical <- my_logical > 0
my_logical

my_letters <- sample(letters[1:26])
my_letters
```


Then, complete the following steps:

  * create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
my_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])
my_list
```

  * use the `typeof()` function to confirm the underlying data types of each component in this list
```{r}
typeof(my_matrix)
typeof(my_logical)
typeof(my_letters)
```

  * combine the underlying elements from the new list into a single atomic vector with the `c()` function.
```{r}
my_list_vector <- c(my_list)
my_list_vector
```

  
  * what is the data type of this vector?  
```{r}
typeof(my_list_vector)
```


#### 3. Create an empty data frame with two variables (= columns) and 26 observations (= rows) below:

  * call the first column/variable `my_unifs` and fill it with 26 random uniform values from 0 to 10
```{r}
my_unifs <- runif(26, min=0, max=10)
my_unifs_DF <- data.frame(my_unifs)
```


  * call the second variable `my_letters` and fill it with 26 capital letters in random order.
```{r}
my_letters <- sample(LETTERS[1:26])
my_letters
```


  * for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with `NA`.
```{r}
my_unifs_DF[sample(nrow(my_unifs_DF), 4),1] <- NA
```


  * for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}

```
  
  
  * re-order the entire data frame to arrange the second variable in alphabetical order.
  
  * calculate the column mean for the first variable.
  
  * rename the second column ‘my_LETTERS’