---
title: "Assignment 4"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---

## Working with matrices, lists, and data frames

1. Assign to the variable `n_dims` a single random integer between 3 and 10.
```{r}
set.seed(111)
n_dims <- sample(3:10, size=1)
print(n_dims)
```
  * Create a vector of consecutive integers from `1` to `n_dims^2^`.
```{r}
n_dims <- 1:n_dims^2
```
  * Use the sample function to randomly reshuffle these values.
```{r}
n_dims <- sample(n_dims)
```
  * create a square matrix with these elements.
```{r}
m <- matrix(data=n_dims, nrow=8)
```
  * print out the matrix.
```{r}
print(m)
```
  * find a function in r to transpose the matrix.
  * print it out again and note how it has changed.
```{r}
t(m)
## matrix was flipped
```
  * calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
mDataframe <- data.frame(m)
rowSums(mDataframe[1,])
rowMeans(mDataframe[1,])

rowSums(mDataframe[8,])
rowMeans(mDataframe[8,])
```

  * set your code up so you can re-run it to create a matrix of a different size by only changing the `n_dims` value
```{r}
n_dims <- sample(3:10, size=1)
n_dims2 <- 1:n_dims^2
n_dims2 <- sample(n_dims2)
m <- matrix(data=n_dims2, nrow=n_dims)
print(m)
t(m)

mDataframe <- data.frame(m)
rowSums(mDataframe[1,])
rowMeans(mDataframe[1,])

rowSums(mDataframe[n_dims,])
rowMeans(mDataframe[n_dims,])
```

